---
title: "R code for Data Science for Beginners"
subtitle: " Day 5: Individual Exercise"
author: "Amanda GJ Fontenot Tuk"  #If multiple, 'c("A", "B")'
date: "2026-9-7"  #r Sys.Date()
format:
  pdf: default
  html: default
---

# clean up your workspace 
```{r}
rm(list=ls(all=TRUE)) # remove all the named objects visible in the environment 
cat("\014") # clean your console
```

# Load the data 
```{r}
library(readr)
library(tidyverse)
library(ggplot2)
 world <- read_csv("world.csv")
```

Remember that Quarto uses a relative path so always save your data in the same folder (or under the same folder) with your Quarto code. 

## Democracy and female representation 

Do democratic countries (`democ_regime == "Yes"`) have better female representation than non-democratic countries? Please answer this question by showing some graphs to assess the relationship between these two variables. 


### Initial data clean up

Analyze y
```{r}
summary(world $ women09)
```

Analyze x
```{r}
summary(world $ democ_regime)
```

Note: There are some missing values. Labels are not intuitive. So, we will deal with these two first.


### Create a smaller data set that omits NA observations.
```{r}
women.dem <- world[ is.na(world $ women09) == FALSE & is.na(world $ democ_regime) == FALSE, ]
```

Let's make sure that we did this correctly. 

```{r}
summary(women.dem $ women09)
summary(women.dem $ democ_regime)
```

We can see that NA cases have been correctly removed. 


### Re-labeling the values
```{r}
women.dem $ dem.dum <- factor(women.dem $ democ_regime, 
                              levels = c("Yes", "No"),
                              labels = c("Democracy", "Autocracy"))
```

## Q1: Describe Y (numerically and graphically) 


### Numerical summary

```{r}
summary(women.dem $ women09)
```

### Graphical summary

```{r}
g <- ggplot(women.dem, aes(x = women09)) + geom_histogram(bins = 30)
g <- g + theme(axis.text.x = element_text(size = 14)) # specify text font axis.text.x to 14
g <- g + xlab("Percentage women in congress") + ylab("Number of countries")
g

```

## Q2: Describe X (numerically and graphically) 


### Numerical summary (frequency table)

```{r}
table(women.dem $ democ_regime)
```

### Graphical summary (bar chart)

```{r}
g2 <- ggplot(women.dem, aes(x = democ_regime)) + geom_bar()
g2 <- g2 + theme(axis.text.x = element_text(size = 12))
g2 <- g2 + xlab("Democratic") + ylab("Number of countries")
g2
```

## Q3: Describe X-Y graphically

### Histograms

```{r}
g3 <- ggplot(women.dem, aes(x = women09)) + geom_histogram()
g3 <- g3 + theme(axis.text.x = element_text(size = 14))
g3 <- g3 + xlab("Percentage women in congress") + ylab("Number of countries")
g3 <- g3 + facet_grid(.~ democ_regime)
g3

```

### Box-plots

```{r}
g4 <- ggplot(women.dem, aes(y = women09, x = democ_regime)) # Remember that we need to specify both y and x here is aes.
g4 <- g4 + geom_boxplot()
g4 <- g4 + theme(axis.text.x = element_text(size = 14))
g4 <- g4 + xlab("Democratic") + ylab("Percentage women in congress")
g4

```