---
title: "R code for Data Science for Beginners"
subtitle: "Day 1: Individual Exercise"
author: "Amanda GJ Fontenot Tuk"  #If multiple, 'c("A", "B")'
date: "2026-9-1"  #r Sys.Date()
format:
  pdf: default
  html: default
---

## Task 0. Getting things ready for programming

If you want to render a pdf as your quarto output, you would first need to download Latex first: 

Install Latex on your computer:
 
* Mac: [here](https://www.youtube.com/watch?v=nUuW_8Hy1qA) </br>
* Windows: [here](https://www.youtube.com/watch?v=kSN-gS4SIqc)


Please read the following links carefully:

* [Getting started with Quarto](https://www.youtube.com/watch?v=_f3latmOhew&t=34s) </br>
* [Markdown Basics](https://rmarkdown.rstudio.com/authoring_basics.html) </br>
* [File naming system](https://style.tidyverse.org/files.html)


## Task 1. Modify the author name above to your name
* Completed. Please view above. 

## Task 2. Perform the following calculations by writing R commands


```{r, message=TRUE, warning=FALSE, echo=TRUE}
### 2-1-Example: Multiply 4 by 30 plus 6, and then raise the answer to the 
###power of 0.5
(4 * (30 +6)) ^ 0.5 # a sample answer is here
```

```{r}
### 2-1-Attempt: Multiply 3 by 20 minus 6, and then raise the answer to the 
###power of 2
(3 *(20-6))^2 #I was not sure if I needed to create an additional equation. 
###I did just in case. 
```

```{r}
### 2-2: Multiply 4 by 30, then add 6 and then raise the answer to the power 
###of 0.5

((4*30)+6)^0.5
```
Note that 2-1 and 2-2 should give you DIFFERENT answers, because these are different operations. 




## Task 3. Working with objects 

### 3-1: Create an object called "X1" which is the number 73

```{r}
### 3-1: Create an object called "X1" which is the number 73
x1=73
x1
```


### 3-2: Create another object called "X2" which is the answer to the sum 99 + 38

```{r}
### 3-2: Create another object called "X2" which is the answer to the sum 99 + 
###38
x2=99+38
x2
```

### 3-3: Multiply X1 and X2 together and store the answer as another object called "X3"

```{r}
### 3-3: Multiply X1 and X2 together and store the answer as another object 
##called "X3"
x3=x1*x2
x3
```

### 3-4: Subtract 1 from X3, and then raise the answer to the power of 0.25

```{r}
### 3-4: Subtract 1 from X3, and then raise the answer to the power of 0.25
(x3-1)^0.25

```

Note: the answer should be 10. 
```{r}
###Note: the answer should be 10.
if((x3-1)^0.25==10){
  print("The Answer is 10.")
}

```
## Task 4. Calculation using objects

Redo the calculation in 2-1 (Multiply 4 by 30 plus 6, and then raise the answer to the power of 0.5), but this time do so by creating objects. 
Take the following steps:


### 4-1: Create an object called part.1 that is equal to "30 plus 6"

```{r}
### 4-1: Create an object called part.1 that is equal to "30 plus 6"
part.1=30+6
part.1
```

### 4-2: Create an object called part.2 that is equal to 4 times part.1
```{r}
### 4-2: Create an object called part.2 that is equal to 4 times part.1
part.2=part.1*4
part.2
```

### 4-3: Raise part.2 to the power of 0.5

```{r}
### 4-3: Raise part.2 to the power of 0.5
answerQuestion4=part.2^0.5
answerQuestion4
```

The numerical answer you get here should be the same as your answer for 2.1

```{r}
if(answerQuestion4==(4 * (30 +6)) ^ 0.5) {
  print("The two answers are the same. 12.")
}
```


## Task 5. Detecting R data types 

**Now, you have seen from the above examples how you should generated your answers and results in `.qmd` and `.rmd` documents. Please continue and finish the following questions.** 

### 5-1 Create a character vector and show that it is character. It'd be nice if you can do it in one line of code

```{r}
### 5-1 Create a character vector and show that it is character. It'd be nice
###if you can do it in one line of code
class(c("Hello", "I","Like","To","Play","Minecraft"))

```

### 5-2 Create a numeric vector and coerce it to a factor vector. It'd be nice if you can do it in one line of code

```{r}
### 5-2 Create a numeric vector and coerce it to a factor vector. It'd be nice 
###if you can do it in one line of code

class(as.factor(c(1,2,3,4,5,6,7,8,9)))

```

## Task 6. Testing NA 

### 6-1 Create a numeric vector with NA 

```{r}
### 6-1 Create a numeric vector with NA 
partSixVector=c(2,4,6,8,9, NA, 15,16,1553)
partSixVector
```

### 6-2 Show me where the NA is located in the vector (using the which() function)

```{r}
### 6-2 Show me where the NA is located in the vector (using the which() 
###function)
which(is.na(partSixVector))

```
### 6-3 Create a numeric vector with NaN  

```{r}
### 6-3 Create a numeric vector with NaN 
partSixVector2=c(2,4,6,8,9, (0/0), 15,16,1553)
partSixVector2
```

### 6-4 Show me where the NaN  is located in the vector (using the which() function)

```{r}
### 6-4 Show me where the NaN  is located in the vector (using the which() 
###function)
which(is.nan(partSixVector2))

```



Finally, execute the entire file. Make sure that you don't get any error messages, such as `"Error: unexpected symbol in ..."` shown in Red. If you receive an error message, it's probably because you forgot to comment out some lines. Comment them out so you won't get any error message. 


